### 2015-10-30 tables/tabNormNull.tex

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 3 & 2 & 1 \\\hline
5.0 & 0.050 & 1.36 & 0 & 1 & 0 \\\hline
2.5 & 0.025 & 1.48 & 0 & 0 & 0 \\\hline
1.0 & 0.010 & 1.63 & 0 & 0 & 0 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
of the null hypothesis for significance levels $\alpha$.
The $c_1$ values were calculated using simulations of normal distributions with $\mu=0$ and $\sigma=1$.
The $c_2$ values were calculated using simulations of normal distributions with $\mu=3$ and $\sigma=2$.
The $c_3$ values were calculated using simulations of normal distributions with $\mu=6$ and $\sigma=3$.
Over all $N_c$ comparisons,
 $\mu(c_1)=0.6979$ and $\sigma(c_1)=0.2339$,
 $\mu(c_2)=0.6838$ and $\sigma(c_2)=0.2697$,
 $\mu(c_3)=0.6809$ and $\sigma(c_3)=0.2220$ .
}
\end{center}
\end{table}