### 2015-11-01 tables/musicGeneral_.tex
melhoras pequenas

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || c | c |}\hline
label & description & events \\\hline\hline
{\bf Pale} & Sanctus 69 from G. P. da Palestrina & 719 \\\hline
{\bf Bach1} & BWV735 from J. S. Bach & 236 \\\hline
{\bf Bach2} & BWV648 from J. S. Bach & 272 \\\hline
{\bf Moza1} & K80 from W. A. Mozart & 538 \\\hline
{\bf Moza2} & K458 from W. A. Mozart & 4218 \\\hline
{\bf Beet1} & Opus 18, n1, mov. 3 from L. van Beethoven & 1289 \\\hline
{\bf Beet2} & Opus 132 from L. van Beethoven & 17884 \\\hline
{\bf Sch\"on} & Opus 19, mov. 2 from A. Sch\"onberg & 102 \\\hline
\end{tabular}
\caption{General description of the music data used for the $c$ values of the next table. Each event is a midi value of a note pitch. Samples where chosen to reflect music history timeline. Works by the same composer were chosen among the first and last 10\% of all he produced.}
\end{center}
\end{table}