### 2015-11-01 tables/osGeneral.tex

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c |}\hline
label & description & events \\\hline
cpu1 & workload of the most active processor & 888 \\\hline
cpu2 & workload of the second most active processor & 422 \\\hline
cpu3 & workload of the third most active processor & 1046 \\\hline
mem & RAM use in kB & 557 \\\hline
p1 & workload use of most consuming process & 1197 \\\hline
m1 & RAM use of most consuming process & 1197 \\\hline
p2 & workload use of second most consuming process & 1197 \\\hline
m2 & RAM use of second most consuming process & 1197 \\\hline
p3 & RAM use of third most consuming process & 1197 \\\hline
m3 & workload use of third most consuming process & 1197 \\\hline
\end{tabular}
\caption{General description of the laptop system status data used for the $c$ values of the next table. Each event is a measure in a snapshot of system status.}
\end{center}
\end{table}