### 2015-11-01 tables/osGeneral_.tex

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || c | c |}\hline
label & description & events \\\hline\hline
{\bf cpu1} & workload of the most active processor & 888 \\\hline
{\bf cpu2} & workload of the second most active processor & 422 \\\hline
{\bf cpu3} & workload of the third most active processor & 1046 \\\hline
{\bf mem} & RAM use in kB & 557 \\\hline
{\bf p1} & workload use of most consuming process & 1197 \\\hline
{\bf m1} & RAM use of most consuming process & 1197 \\\hline
{\bf p2} & workload use of second most consuming process & 1197 \\\hline
{\bf m2} & RAM use of second most consuming process & 1197 \\\hline
{\bf p3} & RAM use of third most consuming process & 1197 \\\hline
{\bf m3} & workload use of third most consuming process & 1197 \\\hline
\end{tabular}
\caption{General description of the laptop system status data used for the $c$ values of the next table. Each event is a measure in a snapshot of system status.}
\end{center}
\end{table}