### 2015-11-04 tables/tabNormNull.tex
adicionadas tabelas com mudanca de amostragem

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ \\\hline
10.0 & 0.100 & 1.22 & 12 & 3 & 11 \\\hline
5.0 & 0.050 & 1.36 & 7 & 0 & 8 \\\hline
2.5 & 0.025 & 1.48 & 1 & 0 & 4 \\\hline
1.0 & 0.010 & 1.63 & 1 & 0 & 1 \\\hline
0.5 & 0.005 & 1.73 & 0 & 0 & 0 \\\hline
0.1 & 0.001 & 1.95 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
        of the null hypothesis for significance levels $\alpha$.
        The $c_1$ values were calculated using simulations of normal distributions with $\mu=0$ and $\sigma=1$.
        The $c_2$ values were calculated using simulations of normal distributions with $\mu=3$ and $\sigma=2$.
        The $c_3$ values were calculated using simulations of normal distributions with $\mu=6$ and $\sigma=3$.
        Over all $N_c$ comparisons,
         $\mu(c_1)=0.8461$ and $\sigma(c_1)=0.2750$,
         $\mu(c_2)=0.7788$ and $\sigma(c_2)=0.2203$,
         $\mu(c_3)=0.8660$ and $\sigma(c_3)=0.2820$ .
        }
\end{center}
\end{table}