### 2015-11-04 tables/textsGeneral.tex
acrescido numero de elementos nas amostras de texto

\begin{table*}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c | c | c | c | c |}\hline
label & description & chars & tokens & sentences & $|kw|$ & $\mu(kw)$ & $\sigma(kw)$ & $|sw|$ & $\mu(sw)$ & $\sigma(sw)$ \\\hline
H,H1,H2 & Hamlet by Shakespeare & 162881 & 37360 & 3106 & 16722 & 3.549 & 1.762 & 9908 & 2.721 & 1.011 \\\hline
B,B1,B2 & King James Version of the Holly Bible & 4332554 & 1010654 & 30103 & 492901 & 3.745 & 1.711 & 289244 & 2.927 & 1.044 \\\hline
M,M1,M2 & Moby Dick by Herman Melville & 1242990 & 260819 & 10059 & 136008 & 4.105 & 2.184 & 75385 & 2.847 & 1.096 \\\hline
E,E1,E2 & Esa\'u e Jac\'o from Machado de Assis & 355706 & 88472 & 3822 & 13984 & 2.186 & 1.376 & 3535 & 1.486 & 0.502 \\\hline
\end{tabular}
\caption{General description of the texts used to exemplify the use of the $c$ statistic.
Individual values of number of characters, tokens, sentences give context.
Mean and standard deviation of the size of known words $kw$ and of the stopwords
$st$ are used in next tables.
Numbers in the labels indicate first and second half of the corresponding text in the next tables.}
\end{center}
\end{table*}