### 2015-11-05 tables/tabWeibullNull_.tex

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l | c | c | c | c | c | c |}\hline
$\alpha N_c$ & $\alpha$ & $c(\alpha)$ & $|C_1(\alpha)|$ & $|C_2(\alpha)|$ & $|C_3(\alpha)|$ & $|C_4(\alpha)|$ \\\hline\hline
10.0 & {\bf 0.100} & 1.22 & 0 & 9 & 5 & 9 \\\hline
5.0 & {\bf 0.050} & 1.36 & 0 & 1 & 3 & 3 \\\hline
2.5 & {\bf 0.025} & 1.48 & 0 & 0 & 1 & 1 \\\hline
1.0 & {\bf 0.010} & 1.63 & 0 & 0 & 1 & 0 \\\hline
0.5 & {\bf 0.005} & 1.73 & 0 & 0 & 1 & 0 \\\hline
0.1 & {\bf 0.001} & 1.95 & 0 & 0 & 0 & 0 \\\hline
\end{tabular}
\caption{The theoretical maximum number $\alpha N_c$ of rejections
        of the null hypothesis for critical values of $\alpha$.
        The $c_1$ values were calculated using simulations of 1-parameter Weibull distributions with $a=0.1$.
        The $c_2$ values were calculated using simulations of 1-parameter Weibull distributions with $a=2$.
        The $c_3$ values were calculated using simulations of 1-parameter Weibull distributions with $a=4$.
        Over all $N_c$ comparisons,
        The $N_o$ values of $c_4$ were calculated using simulations of
         1-parameter Weibull distributions with $a=6$.
        Over all $N_c$ comparisons,
         $\mu(c_1)=0.1107$ and $\sigma(c_1)=0.0652$,
         $\mu(c_2)=0.8079$ and $\sigma(c_2)=0.2417$,
         $\mu(c_3)=0.7775$ and $\sigma(c_3)=0.2404$ .
         $\mu(c_4)=0.8209$ and $\sigma(c_4)=0.2389$ .
        }
\end{center}
\end{table}